# Peer Assessment 1
### Daniel Elbaz
========================================================

## Loading and processing the data set

```{r, echo=TRUE}
data <- read.csv("activity.csv")
data$date <- as.Date(data$date)
```

## Mean and Median of the total number of steps taken each each
### Histogram of Steps Taken
```{r, echo=TRUE}
hist(data$steps)
```

### Mean and Median of Steps
```{r, echo=TRUE}
mean(data$steps, na.rm = TRUE)
median(data$steps, na.rm = TRUE)
```


## Average Daily Activity Pattern
```{r, echo=TRUE}
agg <- aggregate(steps ~ interval, data, mean, na.rm = TRUE)
```
### Plot of Average Daily Activity Pattern by Interval
```{r}
plot(agg[,c("interval","steps")], type = "l")
```

### The Interval with the highest average:
```{r}
agg[which.max(agg$steps),"interval"]
```


## Accounting for NA Values

### Number of NA Values:
```{r, echo=TRUE}
length(data[is.na(data$steps),"steps"])
```

### If the value is NA, use the mean for that interval
```{r, echo=TRUE}
datam <- merge(data, agg, by = "interval")
datam$steps_adj <- ifelse(is.na(datam$steps.x), datam$steps.y, datam$steps.x)
```

## Replot Histogram and Calculate the Mean/median. The mean and median remain the same with this methodology...
```{r, echo=TRUE}
hist(datam$steps_adj)
mean(datam$steps_adj)
median(datam$steps_adj)
```


## Difference in Activity Pattern for Weekdays vs. Weekends.
### Breakout and Plot Average Steps by Weekday/Weekend
```{r, echo=TRUE, message=FALSE}
suppressWarnings(library(ggplot2))
suppressWarnings(library(gridExtra))
data$weekday <- factor(ifelse(weekdays(data$date) %in% c("Saturday","Sunday"),"weekend", "weekday"))
agg2 <- aggregate(steps ~ interval + weekday, data, mean, na.rm = TRUE)
p1 <- ggplot(agg2[agg2$weekday == "weekday",c(1,3)], aes(x = interval, y = steps)) + geom_line() + ggtitle("Weekday")
p2 <- ggplot(agg2[agg2$weekday == "weekend",c(1,3)], aes(x = interval, y = steps)) + geom_line() + ggtitle("Weekend")
grid.arrange(p1, p2, ncol = 1, main = "Average Steps Taken by Interval")
```